### 2025-02-16 src/articles/article10.qmd

# Scholarship

## Scholarship Standards

1. Once per semester, brothers and Candidates will fill out grade release forms.
   a. Grade release forms must be filled out prior to the due date for midterm grades.

1. Brothers on Academic Probation are required to make a grade improvement plan with the Scholarship Chair.
   a. The grade improvement plan will be approved by the Scholarship Chair, the President, and the brother. Failure to follow the grade improvement plan as judged by the Scholarship Chair will result in the brother being brought up for standards.
   a. A brother on Academic Probation is allowed to attend at maximum 3 social events per semester. Their responsibilities for rush are reduced to one third (rounded down) of events. This brother will not be allowed to hold an elected position.

1. Members will be evaluated according to the following criteria:

   a. Any brother with a cumulative Grade Point Average (GPA) of less than 3.0 or an immediately previous semester GPA of less than 3.0 will be placed on Academic Probation until the end of the semester and will be required to attend an Academic Review as described in Article XII, Section K.
      i. The Academic Review must be within the first two weeks of being put on Academic Probation.

   a. If a brother has a midterm progress GPA of less than 3.0, the Scholarship Chair may request the brother to make a grade improvement plan in accordance with Article X, Section A, Item 2a.

   a. If a brother is participating in a program where they do not receive a GPA (e.g. a Co-op) but would otherwise be placed on Academic Probation:
      i. They are not placed on Academic Probation that semester.
      i. The next semester in which they are not participating in such a program, they will be placed on Academic Probation.

## New Member Scholarship

1. During the New Member program, the Scholarship Chair will meet with New Members during a regular New Member meeting to explain the chapter Scholarship Program and answer any questions they have.
1. New Members will be held to the academic standards in Article X, Section A.
1. Candidates will be required to sign the grade releases no later than 14 days of grade releases being made available.
1. A candidate is eligible for initiation if they achieved higher than a 3.0 GPA in the previous semester, or if they have a cumulative GPA of greater than 2.5. Upon receiving a Candidate's GPA, and finding it does not meet requirements, the Marshal shall announce this at the next regularly scheduled chapter meeting.
   a. First semester freshmen are exempt from this rule.
1. Any Candidate that is ineligible to be initiated will have the option to remain a Candidate.
   a. Should a Candidate remain ineligible in any subsequent semester, they may remain a Candidate only with an absolute majority vote by the active chapter.

## Privacy

1. The President, Scholarship Chair, Alumni Advisor, and Faculty Advisor are permitted to know the GPA of each brother.
1. All brothers are permitted to know the names of brothers on Academic Probation. 3. All brothers are permitted to know the names of brothers who cannot hold office due to grades.