### 2025-02-16 src/articles/article4.qmd

# Voting

## Voting Overview

1. All brothers eligible to vote must vote.
1. Only active brothers who are not on probation or financial probation are eligible to vote.
1. Brothers must be present for all of the discussion on a motion in order to be eligible to vote on that motion.
1. A brother not required to pay local dues is ineligible to vote.
1. Unless specified in Article IV, Section B, or explicitly voted on by simple majority, the method of voting will be by a counted show of hands.

## Procedure

1. The following votes will be performed by ballot:
   a. New Members for Initiation
   b. Officer Elections
   c. To overturn an Executive Council (EC) Interpretation of the bylaws (Article I, Section D)
   d. At a suspension trial
   e. To retain a candidate who was not initiated the previous semester

1. Balloting
   a. The Second Guard and the Chaplain will collect and count the ballots.
   b. The chaplain will confirm the count and repeat the results to the presiding officer.
   c. No brother may count the ballots for a matter related to himself.
   d. In the case of absence, conflict of interest, or at the discretion of the chair, the chair may appoint a brother to collect and count the ballots.

1. New Members for Initiation
   a. Balloting on New Members for Initiation will be in accordance with the International Bylaws (IBL) of Theta Chi Fraternity, Article VI, Section 4, paragraph d.
   b. New Members must be approved by a ¾ vote of eligible voting members of the active chapter to be initiated.

1. Officer Elections
   a. Officers will be elected by simple majority vote.
   b. In the case that no nominee receives a simple majority, there will immediately be another vote between all candidates who received the two highest counts of votes.\
   c. Nominations for president and treasurer must be a week before nominations for the other positions in order to make time for CAB interviews.