### 2025-09-29 src/articles/article8.qmd

# Awards {#sec-art-KAF5NVBG}

## Award Election Procedure {#sec-art-WWI418WD}

1. For each award that the chapter gives:
   a. Nominations will be opened prior to voting.
   a. Voting will take place after nominations have been closed.
   a. Voting will be done by show of hands.
   a. A brother may vote *No Confidence* if he is not satisfied with any of the nominees.
   a. There will be no discussion between nominations and voting.
   a. If there are less than 4 nominees, the nominee will be elected to receive the award by {{< var glossary.plurality >}} of voting members present.
   a. Otherwise, there will be a runoff (decided by {{< var glossary.plurality >}}) between the three nominees that received the highest counts of votes.
   a. If a {{< var glossary.plurality >}} of votes cast are *No Confidence*, voting will end and no award will be given.

## Specified Awards {#sec-art-GZYYZN72}

1. Brotherhood Award:
   a. Will be awarded to a current brother who best exemplifies the ideals of Theta Chi Fraternity. b. Will be awarded before the end of the Spring Semester.

1. Alumni Service Award:
   a. Will be given to the alumnus who has done the most for the currently active chapter.
   a. Will be given annually at OX Roast.
   a. A plaque will be maintained by the Alumni Relations Committee Chairman that includes the names of the Alumni who have won the award.

1. Dream Girl Award:
   a. Will be bestowed upon the woman associated with Beta Nu Chapter who has done the most for the presently active chapter.
   a. Will be elected by the second regularly scheduled chapter meeting after initiation during the spring semester.
   a. Voting will be done by {{< var glossary.ballot >}} in direct exception to @sec-art-WWI418WD, Item 1b.

1. Certificate of Appreciation:
   a. May be presented to any person who has demonstrated service of extraordinary care and effort to the chapter.\
   a. Will consist of a certificate of appropriate construction and suitable wording which describes the service for which the certificate is being presented.
   a. Any brother eligible to vote may, at any chapter meeting, nominate any person.
   a. Discussion will be allowed for this award in direct exception to @sec-art-WWI418WD, Item 1c.
   a. A nominee must receive a {{< var glossary.supermajority >}} vote to receive this award in direct exception to @sec-art-WWI418WD, Item 1b.
   a. The chapter may give out any number of this award.

1. The chapter may give out any other awards subject to the procedures defined in @sec-art-WWI418WD.